---
title: "Convert tRNAscan-SE output to GRanges"
author: "Felix G.M. Ernst"
date: "`r Sys.Date()`"
package: tRNAscan2GRanges
abstract: >
  Example conversion of tRNAscan-SE output for sacCer3 to a GRanges object
output: 
  BiocStyle::html_document:
    toc: true
    toc_float: true
    df_print: paged
vignette: >
  %\VignetteIndexEntry{tRNAscan2GRanges}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: references.bib
---

```{r style, echo = FALSE, results = 'asis'}
BiocStyle::markdown(css.files = c('custom.css'))
```

# Getting started

The default tRNAscan-SE [@Lowe1997] output is formatted text document containing
text blocks per tRNA delimited by an empty line.

```{r}
file <- system.file("extdata", 
                    file = "tRNAscan.sort", 
                    package = "tRNAscan2GRanges")

# output for sacCer3
# Before
readLines(con = file, n = 7L)
```

# Converting to GRanges

To access the information in a BioC context the conversion to a GRanges object 
comes to mind. This task is performed by `tRNAscan2GRanges()`, which uses 
regular expressions to extract the information from the text blocks. The result
can be used directly or saved as gff3 file for further use.

```{r}
# After
gr <- tRNAscan2GRanges::tRNAscan2GRanges(file)
head(gr, 2)
```

# Session info

```{r}
sessionInfo()
```

# Literature

Depending on the development on tRNAscan-SE this might become redundant, since
a gff3 export by tRNAscan-SE might resolve the conversion issue. 